Article Title

### 2014-07-14 plos.tex
minor

The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}. Partial correlations figure prominently in probablistic \emph{graphical modelling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}. For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}. More generally, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}. As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functinonal connectivity) between pairs of neurons.

@@

To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Methods). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}). The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).

@@

Orientation tuning was computed by fitting the mean firing rates for each direction of motion $\phi$ using two-peaked von Mises tuning functions $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction. The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of permutations that yielded $R^2$ equal to or higher than that of the unpermuted data.  Cells were considered tuned with $p<0.05$.

@@

To evaluate the estimators, we regularized the conditioned variances by linear shrinkrage toward their mean value across all conditions. This was accomplished by scaling $C_{\sf reg}$ by the diagonal variance adjustment matrix 
\begin{equation}
	Q_c = \delta I + (1-\delta)V_{\sf sample}^{-1} V_{c,\sf sample}
\end{equation}
where $\delta \in [0,1]$ is the variance regularization hyperparameter, which was optimized in the inner loop of cross-validation along with the regularization hyperparameters for $C_{\sf reg}$.

Then the regularized condition-specific covariance matrices became
\begin{equation}
	C_{c,\sf reg} = Q_c^{\frac 1 2} C_{\sf reg} Q_c^{\frac 1 2} 
\end{equation}

@@

%\section*{Figure Legends}
\newpage

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./figures/Supp01.pdf}
\end{center}
\caption{
	{\bf Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.}  
	{\bf A.} The cross-validated loss function (Eq.~\ref{eq:full-loss}) of the data from the example as a function of the values of hyperparameters $\alpha$ and $\beta$ of the minimization problem in Eq.~\ref{eq:ma}. In all panels, the red cross marks the optimal value found by the pattern search algorithm described in Methods.
	{\bf B.} The connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of $\alpha$ and $\beta$ for the example site.
	{\bf C.} The number of latent units or the rank of the low-rank component $L$ of the inverse covariance (Eq.~\ref{eq:c-sl}) as a function of hyperparameters $\alpha$ and $\beta$ for the example site.
	{\bf D.} The loss function as a function of the the connectivity and the number of latent units for the exapmle site.
}
\end{figure}

>>>>

The \emph{partial correlation} between a pair of variables is the Pearson correlation coefficient of the residuals of the linear least-squares predictor of their activity based on all the other variables, excluding the pair \cite{Anderson:2003, Whittaker:1990}. Partial correlations figure prominently in probabilistic \emph{graphical modeling} wherein the joint distribution is explained by sets of two-way interactions \cite{Whittaker:1990}. For the multivariate Gaussian distribution, zero partial correlations indicate conditional independence of the pair, implying a lack of direct interaction \cite{Dempster:1972, Whittaker:1990}. More generally, partial correlations can serve as a measure of conditional independence under the assumption that most dependencies in the system include strong linear effects \cite{Whittaker:1990,Baba:2004}. As neural recordings become increasingly dense, partial correlations may prove useful as indicators of conditional independence (lack of functional connectivity) between pairs of neurons.

@@

To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}). The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).

@@

Orientation tuning was computed by fitting the mean firing rates for each direction of motion $\phi$ using two-peaked von Mises tuning functions $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction. The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of permutations that yielded $R^2$ equal to or higher than that of the original data.  Cells were considered tuned with $p<0.05$.

@@

To evaluate the estimators, we regularized the conditioned variances by linear shrinkage toward their mean value across all conditions. This was done by scaling $C_{\sf reg}$ by the conditioned variance adjustment matrix $Q_c = \delta I + (1-\delta)V_{\sf sample}^{-1} V_{c,\sf sample}$ to produce the conditioned regularized covariance estimate:
\begin{equation}
	C_{c,\sf reg} = Q_c^{\frac 1 2} C_{\sf reg} Q_c^{\frac 1 2},
\end{equation}
The variance regularization parameter $\delta \in [0,1]$ was optimized in the inner loop of cross-validation along with the regularization hyperparameters for $C_{\sf reg}$.

@@

\section*{Figure Legends}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./figures/Supp01.pdf}
\end{center}
\caption{
	{\bf Optimization of hyperparameters of the $C_{\sf sparse+latent}$ estimator.}  
	{\bf A.} Validation loss (Eq.~\ref{eq:full-loss}) for the example site in \figref{2} and \figref{4} as a function of the hyperparameters $\alpha$ and $\beta$ of the $C_{\sf sparse+latent}$ estimator (Eq.~\ref{eq:c-sl} and Eq.~\ref{eq:ma}). In all panels, the red cross marks the optimal value found by the pattern search algorithm described in Methods.
	{\bf B.} The connectivity ($1-\mbox{sparsity}$) of the sparse component $S$ as a function of $\alpha$ and $\beta$ for the example site.
	{\bf C.} The number of latent units, \ie the rank of the low-rank component $L$ as a function of hyperparameters $\alpha$ and $\beta$.
	{\bf D.} The loss function as a function of the connectivity and the number of latent units.
}
\end{figure}
